### 2020-11-28 latex/acrobot.tex
Completed the edit of oscillation proof, ready to send to Manfredi

A small enough fixed value \(I > 0\) will enable the acrobot to 
escape any compact subset of \(\mathcal{O}\) in finite time. 
What's more, the acrobot's phase will flow towards the heteroclinic orbit
corresponding to \(E(\pi,0)\), as in Figure
\ref{fig:acrobot-oscillation-domain}.
Hence, the first result of Theorem \ref{thm:acrobot-energy-stabilization} states
that all acrobots constrained by ~\eqref{eqn:acrobot-constraint} 
will gain enough energy to begin rotating around the bar.
In the worst case, the acrobot will at least perform a swing-up routine to reach
the \((q_u,p_u) = (\pi,0)\) equilibrium.

If we want the acrobot to achieve giants with energy
\(E(0,\bar{\rho})\), the acrobot needs to satisfy the assumption on the
integral of \(b(\beta,\rho_0)\).
In this case, orbits will escape compact subsets of \(\Omega\) in finite time
until the acrobot has reached a momentum of at least \(\bar{\rho}\).
One example of this behaviour is displayed in Figure \ref{fig:acrobot-omega}.
Note that the value of \(I\) will depend on \(\bar{\rho}\).

\begin{figure}
    \centering
    % TODO: show an image of two subfigures: one is O, with a spiraling
    % orbit heading towards the heteroclinic orbit E(pi,0).
    % The second is Omega, with a spiral leaving O and becoming
    % a set of rotations that approaches E(0,rho_bar)
    \begin{subfigure}[t]{0.45\textwidth}
        % TODO: 
        %includestandalone[]{images/O_epsilon}
        \caption{The set \(\mathcal{O}\). Orbits starting in this set
            will reach the level set \(E(\pi,0)\) of the nominal pendulum in
            finite time.}
        \label{fig:acrobot-oscillation-domain}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        %TODO:
        %includestandalone[]{images/Omega_epsilon}
        \caption{The set \(\Omega\). Orbits starting in this set will
            reach the level set \(E(0,\bar{\rho})\) of the nominal pendulum in
            finite time.}
            \label{fig:acrobot-omega}
    \end{subfigure}
    \caption{The sets defined by Theorem \ref{thm:acrobot-energy-stabilization}
        on which the acrobot gains energy.}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        % TODO: figure of acrobot with I = 0
        %\includestandalone[]{images/acrobot_I0}
        \caption{A simple pendulum with two masses.}
        \label{fig:acrobot-I0}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        % TODO: figure showing oscillations and rotations as level curves of E
        %\includestandalone[]{images/pendulum_level_sets}
        \caption{Level sets of the pendulum energy function.
            The blue ellipse represents oscillations; the red curve
            represents rotations; and the black ellipse is the heteroclinic
            orbit with energy \(E(\pi,0)\) forming the boundary of the
            oscillation and rotation regions.}
        \label{fig:pendulum-level-sets}
    \end{subfigure}
    \caption{Our constrained acrobot is a simple pendulum when \(I = 0\).}
\end{figure}

@@

\begin{figure}
    \centering
    % TODO: Add a figure with one oscillation mark at a distance mu from
    % the origin to the intersection on the q-axis. We also draw the boundary of
    % mathcalO and shade the interior, and draw a line to the oscillation with
    % clockwise angle alpha
    %\includestandalone[]{images/mu_intersection}
    \caption{The domain \(\mathcal{O}\) where a pendulum oscillates.
    The pseudo-radius \(\mu\) corresponds to the
    intersection of an orbit of oscillation with the \(q_u\)-axis.
    The psuedo-angle \(\alpha\) is taken to be clockwise positive because 
    oscillations of a pendulum move clockwise on \(\mathcal{O}\).}
    \label{fig:mu-intersection}
\end{figure}

@@

%Suppose at time \(t = 0\) we have initialized the acrobot at \((\alpha,\mu) = (0,\mu_0)\).
%Since \(\alpha\) is continuously increasing, we know that the orbit will do one
%of two things: it will either leave \(\mathcal{O}\) so that the acrobot is now rotating,
%or it will remain in \(\mathcal{O}\) and hit the \(q_u\)-axis at some point 
%\((\alpha,\mu) = (\pi,q_1)\).
%From there, the acrobot could escape \(\mathcal{O}\) along the second half of the orbit or
%return to some point \((\alpha,\mu) = (0,\tilde{u})\) (as illustrated in Figure
%\ref{fig:acrobot-possible-orbits}).
%If \(\tilde{u} < \mu_0\) then the acrobot has lost energy; 
%if \(\tilde{u} = \mu_0\), the acrobot is on a closed orbit;
%and if \(\tilde{u} > \mu_0\) the acrobot has gained energy.
%Which situation occurs might depend entirely on the choice of
%\(\mu_0\).
%
%\begin{figure}
%    \centering 
%    % TODO: Draw the domain mathcalO, an initial point (q0,p0) = (mu0,0), and show the
%    % half orbit hitting the left q-axis at (-m,0).
%    % Then show the 
%    % 4 clockwise half orbits starting at (-m,0): 
%    % one orbit exits mathcalO on the top right corner;
%    % one orbit hits the right q-axis at (mu0+x,0); 
%    % one hits (mu0,0);
%    % and one hits (mu0-x,0) for some x > 0.
%    %includestandalone[width=0.5\textwdith]{images/acrobot_possible_orbits}
%    \caption{If the acrobot is initialized at \((q_u,p_u) = (\mu_0,0)\), then
%        either the orbit escapes \(\mathcal{O}\) into the rotation
%        zone (purple), or the orbit returns to the \(q_u\)-axis 
%        with a higher (red), equal (black), or lower (blue) pseudo-radius than when it
%        started.}
%    \label{fig:acrobot-possible-orbits}
%\end{figure}

@@

We now take a first-order Taylor approximation of \(\hat{\mu}(\alpha,\mu_0,I)\)
around the nominal system \(\hat{\mu}_0(\alpha,\mu_0)\), which 
expands into 
\begin{equation}\label{eqn:acrobot-muhat-approx}
    \hat{\mu}(\alpha,\mu_0,I) = \hat{\mu}_0(\alpha,\mu_0) + I
    \hat{\mu}_1(\alpha,\mu_0)
    + R(\alpha,\mu_0,I)
    ,
\end{equation}
where \(R(\alpha,\mu_0,I)\) is smooth and \(O(I^2)\).

@@

\begin{figure}
    \centering
    % TODO: The plot of Q(mu) from MATLAB
    %\includegraphics[width=0.5\textwidth]{images/Qmu.png}
    \caption{The plot of \(Q(\mu_0)\).}
    \label{fig:acrobot-Q}
\end{figure}

@@

Show energy gain on \(\mathcal{O}\) is equivalent to showing \(\mu(n)\)
eventually reaches \(\pi\).
That is, we want to find \(I^\star \in ]0, I_1]\) where, for all 
\(\delta > 0\) and \(\mu_0 \in ]0,\pi-\delta]\), there exists \(N > 0\) so that
for all \(n \geq N\), \(\mu(n) \notin [0,\pi-\delta]\).
A sufficient condition for this characterization is to show 
\(P_\mathcal{O}(\mu_0) > \mu_0\) for all \(\mu_0 \in ]0,\pi[\),

@@

We have shown \(0\) is a repeller of the discrete time system, which means
there exists some (unknown) \(\epsilon > 0\) where the interval
\(]0,\epsilon[\) is negatively invariant for ~\eqref{eqn:muhat-discrete}.
What's more, all solutions starting in this interval will flow towards the value
\(\mu = \epsilon\).
In other words, we have found a value \(I_2 > 0\) small enough
that \(\mu(n)\) is strictly increasing on \(]0,\epsilon[\).

Next, recall that \(R(\pi,\mu_0,I)\) is smooth in all its parameters as well as
being \(O(I^2)\).
Therefore, it is bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
by some value \(R(\pi,\underbar{\mu},I)\). 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some
\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in [-I_3, I_3]\),
\[
    R(\pi,\mu_0,I) \geq R(\pi, \underbar{\mu}, I) > -I^2 r
    .
\]
Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
for\(\mu_0 \in [\epsilon,\pi[\) we have
\[
    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
    .
\]
There exists a small enough \(I^\star \in ]0,I_3]\) making
\(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\), which means
\[
    P_\mathcal{O}(\mu_0) > \mu_0
    ,
\]
for all \(\mu_0 \in [\epsilon,\pi[\).

We have thus shown that \(P_\mathcal{O}(\mu_0) > \mu_0\) for all
\(\mu_0 \in ]0,\pi[\).
Looking at ~\eqref{eqn:muhat-discrete}, this means
\(\mu(n+1) > \mu(n)\) and all solutions of the discrete time system will flow
towards \(\mu = \pi\).
Hence, orbits of the constrained dynamics will eventually escape
\(\mathcal{O}\).

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), 
there exists \(I > 0\) small enough that the constraint 
~\eqref{eqn:acrobot-constraint} is injecting energy on \(\mathcal{O}\).

Setting \(I < 0\), the Poincar\'{e} section now satisfies
\[
    P_\mathcal{O}(\mu_0) < \mu_0
    ,
\]
thanks to the same arguments presented in this section. 
The constraint is therefore dissipating energy on \(\mathcal{O}\).

>>>>

A small enough fixed value \(I > 0\) will enable the acrobot to 
escape any compact subset of \(\mathcal{O}\) in finite time. 
What's more, the acrobot's phase will flow towards the homoclinic orbit
corresponding to \(E(\pi,0)\), as in Figure
\ref{fig:acrobot-oscillation-domain}.
Hence, the first result of Theorem \ref{thm:acrobot-energy-stabilization} states
that all acrobots constrained by ~\eqref{eqn:acrobot-constraint} 
will gain enough energy to begin rotating around the bar.
In the worst case, the acrobot will at least perform a swing-up routine to reach
the \((q_u,p_u) = (\pi,0)\) equilibrium.

If we want the acrobot to achieve giants with energy
\(E(0,\bar{\rho})\), the acrobot needs to satisfy the assumption on the
integral of \(b(\beta,\rho_0)\).
In this case, orbits will escape compact subsets of \(\Omega\) in finite time
until the acrobot eventually reaches a momentum of at least \(\bar{\rho}\).
One example of this behaviour is displayed in Figure \ref{fig:acrobot-omega}.
Note that the value of \(I\) will depend on \(\bar{\rho}\).

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \includestandalone[width=\linewidth]{images/acrobot_oscillation_domain}
        \caption{The set \(\mathcal{O}\). An orbit starting in this set (blue)
            will reach the level set \(E(\pi,0)\) of the nominal pendulum in
            finite time.}
        \label{fig:acrobot-oscillation-domain}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \includestandalone[width=\linewidth]{images/acrobot_omega}
        \caption{The set \(\Omega\). Orbits starting in this set will
            reach the level set \(E(0,\bar{\rho})\) of the nominal pendulum in
            finite time.}
            \label{fig:acrobot-omega}
    \end{subfigure}
    \caption{The sets defined by Theorem \ref{thm:acrobot-energy-stabilization}
        on which the acrobot gains energy.}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \includestandalone[]{images/acrobot_I_zero}
        \caption{A simple pendulum with two masses.}
        \label{fig:acrobot-I0}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.5\textwidth}
        \includestandalone[]{images/pendulum_level_sets}
        \caption{Level sets of \(E\).
            The blue ellipse is an oscillation; the red curves
            are rotations; and the black ellipse is the homoclinic
            orbit with energy \(E(\pi,0)\).}
        \label{fig:pendulum-level-sets}
    \end{subfigure}
    \caption{Our constrained acrobot is a simple pendulum when \(I = 0\).}
\end{figure}

@@

\begin{figure}
    \centering
    \includestandalone[width=0.5\textwidth]{images/mu_intersection}
    \caption{The domain \(\mathcal{O}\) (blue) where a pendulum oscillates.
    The pseudo-radius \(\mu\) corresponds to the
    intersection of an oscillation (red) with the \(q_u\)-axis.
    The psuedo-angle \(\alpha\) is taken to be clockwise positive because 
    oscillations of a pendulum move clockwise on \(\mathcal{O}\).}
    \label{fig:mu-intersection}
\end{figure}

@@

@@

We now take a first-order Taylor approximation of \(\hat{\mu}(\alpha,\mu_0,I)\)
around the nominal system \(\hat{\mu}_0(\alpha,\mu_0)\), which 
expands into 
\begin{equation}\label{eqn:acrobot-muhat-approx}
    \hat{\mu}(\alpha,\mu_0,I) = \hat{\mu}_0(\alpha,\mu_0) + I
    \hat{\mu}_1(\alpha,\mu_0)
    + R(\alpha,\mu_0,I)
    .
\end{equation}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{images/Qmu.png}
    \caption{The plot of \(Q(\mu_0)\).}
    \label{fig:acrobot-Q}
\end{figure}

@@

Proving the constrained dynamics gain energy on \(\mathcal{O}\) is equivalent to
showing \(\mu(n)\) eventually reaches \(\pi\).
That is, we want to find \(I^\star \in\, ]0, I_1]\) where, for all 
\(\delta > 0\) and \(\mu_0 \in ]0,\pi-\delta]\), there exists \(N > 0\) so that
for all \(n \geq N\), \(\mu(n) \notin [0,\pi-\delta]\).
A sufficient condition for this characterization is to show 
\(P_\mathcal{O}(\mu_0) > \mu_0\) for all \(\mu_0 \in\, ]0,\pi[\),

@@

We have shown \(0\) is a repeller of the discrete time system, which means
there exists some (unknown) \(\epsilon > 0\) where the interval
\(]0,\epsilon[\) is negatively invariant for ~\eqref{eqn:muhat-discrete}.
What's more, all solutions starting in this interval will flow towards the value
\(\mu = \epsilon\), so we must have \(P_\mathcal{O}(\mu_0) > \mu_0\) for all 
\(\mu_0 \in \, ]0,\epsilon[\).

Next, recall that \(R(\pi,\mu_0,I)\) is smooth in all its parameters as well as
being \(O(I^2)\).
Therefore, it is bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
by some value \(R(\pi,\underbar{\mu},I)\). 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some
\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in [-I_3, I_3]\),
\[
    R(\pi,\mu_0,I) \geq R(\pi, \underbar{\mu}, I) > -I^2 r
    .
\]
Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
for\(\mu_0 \in [\epsilon,\pi[\) we have
\[
    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
    .
\]
Choosing \(I^\star \in\, ]0,I_3]\) so that
\(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\)
means that
\( P_\mathcal{O}(\mu_0) > \mu_0\)
for all \(\mu_0 \in [\epsilon,\pi[\).

We have thus shown that \(P_\mathcal{O}(\mu_0) > \mu_0\) for all
\(\mu_0 \in ]0,\pi[\).
Looking at ~\eqref{eqn:muhat-discrete}, this means
\(\mu(n+1) > \mu(n)\) and all solutions of the discrete time system will flow
towards \(\mu = \pi\).
Hence, orbits of the constrained dynamics will eventually escape
\(\mathcal{O}\).
We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), 
there exists \(I > 0\) small enough that the constraint 
~\eqref{eqn:acrobot-constraint} is injecting energy on \(\mathcal{O}\).

By the same arguments presented in this section, the Poincar\'{e} section
satisfies \( P_\mathcal{O}(\mu_0) < \mu_0 \) when \(I < 0\). 
In this case, the constraint dissipates energy on \(\mathcal{O}\).
